### 2023-07-18 Rscripts/AMPHIBIAN.DATA.Rmd
hi this is the first commit! were uploading all the old codebase

---
title: "AMPHIBIAN SURVEYS"
author: "RAFAEL BAEZ"
date: "2022-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#dependencies
library(googlesheets4)
library(tidyverse)
library(vegan)
library(viridisLite)

#load in data sheet

```

## Project report

```{r data set up}
#load in Ohio data
MAIN<-read_sheet("https://docs.google.com/spreadsheets/d/1VIVJd4BpC5KS2o7APdG4W5Sf2H4hGTDMdPZ6oAZEL7E/edit#gid=0") %>% 
  arrange(Site)  %>%
  mutate(Beavers = as.logical(Beavers),
         Fish = as.logical(Fish),
         AMD = as.logical(Acid_Mine_Drainage)) %>%
  select(!Acid_Mine_Drainage)

#check structure of data sheet
str(MAIN)

```

```{r}
#DEMO: How to vegan package
###Diversity calculations----
#arrange data by site
SpeciesData <- MAIN %>%
  select(Site, Number_GreenFrogs, Number_AmericanBullfrogs, Number_EasternNewts, Number_AmericanToads, Number_NorthernSlimySalamander, Number_RavineSalamander, Number_GrayTreeFrog)%>%
  arrange(Site)

#make dataframe of species abundance data
SD <- SpeciesData %>%
  select(Number_GreenFrogs, Number_AmericanBullfrogs, Number_EasternNewts, Number_AmericanToads, Number_NorthernSlimySalamander, Number_RavineSalamander, Number_GrayTreeFrog)

#make abscence prescence table, calculate number of taxa per site
AbsPres <- SpeciesData %>%
  transmute(GrFr = case_when(Number_GreenFrogs > 0 ~ 1, TRUE ~ 0), 
     AmBu = case_when(Number_AmericanBullfrogs > 0 ~ 1, TRUE ~ 0), 
     EaNe = case_when(Number_EasternNewts > 0 ~ 1, TRUE ~ 0),
     AmTo = case_when(Number_AmericanToads > 0 ~ 1, TRUE ~ 0),
     NSlSa = case_when( Number_NorthernSlimySalamander > 0 ~ 1, TRUE ~ 0),
     RaSa = case_when(Number_RavineSalamander > 0 ~ 1, TRUE ~ 0),
     GrTrFr = case_when(Number_GrayTreeFrog > 0 ~ 1, TRUE ~ 0))

#it has the diversity now but i have to link it back to main to run my tests
MAIN<-MAIN%>% mutate("ShannonDiversity" =diversity(SD),
                     "SimpsonDiversity" = diversity(SD, index = 'simpson'),
                     "n_taxa" = AbsPres$n_taxa) %>%
  mutate("evenness" = ShannonDiversity/log(Richness))

MAIN
#now my main data frame has diversity at the end
#i must add values for richness 
#add values for average population size

###summary statistics----
#summary table of mean species counts and se species counts
Pop_summ <- SpeciesData %>%
  mutate(se_GF = sd(Number_GreenFrogs)/sqrt(8),
         se_AB = sd(Number_AmericanBullfrogs)/sqrt(8),
         se_EN = sd(Number_EasternNewts)/sqrt(8),
         se_AT = sd(Number_AmericanToads)/sqrt(8),
         se_NSS = sd(Number_NorthernSlimySalamander)/sqrt(8),
         se_RS = sd(Number_RavineSalamander)/sqrt(8),
         se_GTF = sd(Number_GrayTreeFrog)/sqrt(8)) %>%
  summarise(GF = mean(Number_GreenFrogs), se_GF = mean(se_GF), #calculate means
            AB =mean(Number_AmericanBullfrogs), se_AB = mean(se_AB),
            EN = mean(Number_EasternNewts), se_EN = mean(se_EN),
            AT = mean(Number_AmericanToads), se_AT = mean(se_AT),
            NSS = mean(Number_NorthernSlimySalamander), se_NSS = mean(se_NSS),
            RS = mean(Number_RavineSalamander), se_RS = mean(se_RS),
            GTF = mean(Number_GrayTreeFrog), se_GTF = mean(se_GTF))

str(Pop_summ)
#pivot summary stat table for graphing
#species order 
Pop_summ1 <- pivot_longer(Pop_summ, 
             cols = c(GF,AB,EN,AT,NSS, RS,GTF), names_to = "species")

```

```{r}
#diversity plots----
#shannon diversity index

#boxplots----

#linear model
Shannon_anova <- aov(data = MAIN, ShannonDiversity ~ Fish)
anova(Shannon_anova)

Simpson_aov <- aov(data = MAIN, SimpsonDiversity ~ Fish)
anova(Simpson_aov)
```

```{r}
#metamorphosis data
#green frogs as example
#calculate  post metamorphosis prop
MAIN <- MAIN %>%
  mutate(GF_pm_prop = Number_GreenFrogs_PostMetamorphosis/Number_GreenFrogs,
          AB_pm_prop = Number_AmericanBullfrogs_Postmetamorphosis/Number_AmericanBullfrogs,
         EN_pm_prop = Number_EasternNewts_PostMetamorphosis/Number_EasternNewts)

#plot of props
#green frogs
ggplot(MAIN, aes(x = AMD, y = GF_pm_prop, fill = AMD)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(height = 0, width = 0.3, shape = 21, stroke = 0.7)

#American Bullfrogs
ggplot(MAIN, aes(x = AMD, y = AB_pm_prop, fill = AMD)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(height = 0, width = 0.3, shape = 21, stroke = 0.7)

#eastern Newts
ggplot(MAIN, aes(x = AMD, y = EN_pm_prop, fill = AMD)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(height = 0, width = 0.3, shape = 21, stroke = 0.7)

#stats tests
prop_test <- MAIN %>%
  select(AMD, Fish, Beavers, GF_pm_prop, AB_pm_prop, EN_pm_prop)

#green frog 
GF <- prop_test %>%
  select(AMD, GF_pm_prop) %>%
  mutate(rn = row_number()) %>%
  pivot_wider( names_from = AMD, values_from = GF_pm_prop)
AMD_T_GF <- na.omit(c(GF$`TRUE`))
AMD_F_GF <- na.omit(c(GF$`FALSE`))
t.test(x = AMD_T_GF, y = AMD_F_GF)

#American Bullfrog
AB <- prop_test %>%
  select(AMD, AB_pm_prop) %>%
  mutate(rn = row_number()) %>%
  pivot_wider( names_from = AMD, values_from = AB_pm_prop)
AMD_T_AB <- na.omit(c(AB$`TRUE`))
AMD_F_AB <- na.omit(c(AB$`FALSE`))
t.test(x = AMD_T_AB, y = AMD_F_AB)

```

![soda.](Photos/unnamed.jpg)