### 2023-07-24 Rdocs/ManuscriptLPP.Rmd
deleted defunct data and folder added a markdown file for the manuscript repurposed old controller script as stats script added some dependencies to plot_source

---
title: Biodegradable alternatives to plastic mesh bags conventionally used in freshwater
  macroinvertebrate analysis within the Bronx River, NYC
author: "R.E. Segui-Baez, G.A. Rueda Moreno, R. Ward, K. Schneider Paolantonio"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
#knitr options----
knitr::opts_chunk$set(echo = TRUE)
#dependencies----
require(GGally)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r pre-requisite scripts, eval =FALSE}
#source data processor scripts----
source("Rscripts/biodiversity.R")#script to process biodiversity data
source("Rscripts/water_quality.R")#script to process water quality data
source("Rscripts/statistics.R") #script to analyze previously loaded data
source("Rscripts/plot_source.R") #source code for custom plotting functions

```

## Including Plots

You can also embed plots, for example:

```{r plots, echo=FALSE}
ggpairs(select(bd_list$biodiversity, MidgeFlies:Planaria))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.