### 2024-02-22 Rdocs/Plots.Rmd
Fixed the clusterfuck that was waterquality data, spaghetti code abounds
now The water quality plots are actually sensible, and thankfully stats remain unchanged.

sqrt transformed Nitrogen and Phosphorous, and now they are normal so I added them to the normal ANOVA group

Also, switched from doing independence test with coin and now im doing kruskal  wallis tests twice, once for site and once for date (not statistically responsible i know, but what choice do we have????)

I am dreading actually writing the manuscript

---
output:
  pdf_document: default
  html_document: default
---

```{r pre-requisite scripts, echo=FALSE, results='hide',message=FALSE,warning=FALSE}
#set knitr options----
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
#source data processor scripts----
source("Rscripts/biodiversity.R")#script to process biodiversity data
source("Rscripts/water_quality.R")#script to process water quality data
source("Rscripts/stats.R") #script to analyze previously loaded data

```

```{r setup,echo=FALSE, results='hide',message=FALSE,warning=FALSE}
#dependencies----
require(tidyverse)
require(GGally)
require(vegan)
require(ggpubr)
require(tidyverse)
require(viridisLite)
require(rlang)
require(car)
require(corrplot)
require(factoextra)

#plotting source code----
#set plotting theme function 
theme_rafa <- function(base_size = 12){
  theme_pubr(base_family="sans") %+replace%
    theme(panel.background  = element_blank(),
          plot.background = element_rect(fill="transparent", colour=NA), 
          legend.background = element_rect(fill="transparent", colour=NA),
          legend.key = element_rect(fill="transparent", colour=NA),
          legend.text = element_text(size=base_size * 0.7),
          legend.title = element_text(size = base_size * 0.8, 
                                      face = "bold"),
          axis.text = element_text(size = base_size*0.85),
          axis.title = element_text(size = base_size * 1.2),
          plot.title.position = "panel",
          plot.caption.position = "panel",
          plot.title = element_text(
            size = base_size * 1.25,
            face = "bold", 
            hjust = 0.5,
            vjust = 1),
          plot.subtitle = element_text(
            size = base_size * 0.9,
            vjust = 1),
          plot.caption =  element_text(
            size = base_size*0.75,
            hjust = 0.5)
    )
}

#bag type barplot function:
#PARAMS: 
  #index(string): a string specifying the desired index
bt_barplot <- function(index){
  #tukey tibble
  groupings_tibble <- diversity_tukeys[[index]]$groups %>%
rownames_to_column("material")
  #calculate mean and standard error of chosen index
  new_tbl <- LPP_FullData%>%
  filter(is.na(material) == FALSE)%>% #remove na values from material before calculations
    group_by(material) %>% #group by material
    mutate(index_se = sd(get(index), na.rm = T)/sqrt(n())) %>% # calc SE
    summarise(index = mean(get(index), na.rm = T), index_se = mean(index_se)) #summarise into mean
  #merge two tibbles
  new_tbl <- inner_join(new_tbl, groupings_tibble, by ="material")

  #plotting
  ggplot(new_tbl, aes(x = reorder(material, -index, sum), y = index)) +
    geom_col(aes(fill = reorder(material, -index, sum))) +
    geom_errorbar(aes(ymin = index - 2*index_se, ymax = index + 2*index_se)) +
geom_text(aes(y = (index + 3*index_se), label =groups)) +

    theme_bw() +
    scale_fill_viridis_d()+
    theme_rafa(base_size = 15) +
    theme(axis.text.x = element_text(angle = -15, vjust = 0.5))+
    labs(x = "material", title = "Index by Bag material", y = paste(index),
         fill = "material")
}

#water quality time series plots by site
#PARAMS:
  #param(string): a string specifying the water quality parameter
wq_ts_BySite <- function(param){
  #create symbol of parameter name and se for tidy eval in ggplot2 call
  param_se <- sym(paste(param,"se", sep = "_"))
  param<- sym(paste(param, "mean", sep = "_"))

  ggplot(water_quality, aes(x = date, y = !!param, color = site)) +
    #geom layers
    geom_errorbar(aes(ymin = !!param - 2*!!param_se, ymax = !!param +2*!!param_se), color = "black") +
    geom_point( size = 2) +
    facet_wrap(~site) +
    #labels
    labs(title = paste(param, "Seasonality"), x = "Date")+
    #theme adjustments
    theme_bw() +
    theme_rafa(base_size = 15) +
    scale_color_viridis_d()+
  scale_fill_viridis_d()
}

wq_boxplot_BySite <- function(param){
  #create symbol of parameter name and se for tidy eval in ggplot2 call

  param_se <- sym(paste(param,"se", sep = "_"))
  param<- sym(paste(param, "mean", sep = "_"))
  ggplot(LPP_FullData, aes(x = site, y = !!param, fill = site)) +
    #geom layers
    geom_boxplot()+
    geom_jitter(width = 0.5, height = 0.0, shape = 21, size = 2) +
    #labels
    labs(title = paste(param, "across sites"), x = "Site")+
    #theme adjustments
    theme_bw() +
    theme_rafa(base_size = 15) +
    scale_fill_viridis_d(option = "E",begin = 0.1, end = 0.9)
}
```

# Leaf Pack Project Plots

```{r wq plots, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
#water quality dot plots----
#water quality by date on the y and site as color
wq_ts_BySite("P") +labs( title = "Phosphorous Seasonality",y = "Phosphorous (mg/L)")
wq_ts_BySite("pH")
wq_ts_BySite("NH3") +labs(y = "Ammonia (mg/L)")
wq_ts_BySite("DO") + labs(title = "Dissolved Oxygen Seasonality",
                          y = "% DO")
wq_ts_BySite("Temp") + labs(title = "Temperature Seasonality",
                            y = "Temp (°C)")
wq_ts_BySite("Cond") + labs(title = "Conductivity Seasonality", y = "Conductivity (mS)")
wq_ts_BySite("Nitrogen") + labs(y = "Nitrogen (mg/L)")
wq_ts_BySite("Flow") + labs(title = "Flow rate Seasonality",
                            x = "Date", y = "Flow (m/s)")

wq_boxplot_BySite("P") + labs(y = "Phosphorous (mg/L)", title = "Phosphorous Variation Across Sites")
wq_boxplot_BySite("pH")
wq_boxplot_BySite("NH3")+ labs(y = "Ammonia (mg/L)", title = "Ammonia Variation Across Sites")
wq_boxplot_BySite("Nitrogen")+ labs(y = "Nitrogen (mg/L)", title = "Nitrogen Variation Across Sites")
wq_boxplot_BySite("DO") + labs(y = " % DO")
wq_boxplot_BySite("Flow") + labs(y = "Flow (m/s)", title = "Flow Rate Variation Across Sites")
wq_boxplot_BySite("Cond") + labs(y = "Conductivity (mS)", title = "Conductivity Variation Across Sites")
wq_boxplot_BySite("Temp") + labs(y = "Temp (°C)", title = "Temperature Variation Across Sites")

#biodiversity plots----
bt_barplot( "richness")+ labs(x = "Material", y="Richness")  
bt_barplot( "shannon")+ labs(x = "Material", y= "Shannon Diversity")
bt_barplot( "simpson")+ labs(x = "Material",y= "Simpson Diversity")
bt_barplot( "invsimp")+ labs(x = "Material",y= "Inverse-Simpson Diversity")
bt_barplot("even")+ labs(x = "Material",y= "Shannon Evenness")

#Bray Curtis Dissimil plots----
#by site
ggplot(BC_BySite_tibble, aes(x = site_comp, y = bc_score, fill = site_comp)) +geom_col() +
  theme_rafa() + coord_cartesian(ylim = c(0, 1)) + scale_fill_viridis_d() +
  labs(title = "Community Dissimilarity between sites",
       x = "Site Comparisons", y =" Bray Curtis Dissimilarity", fill = "Comparisons")

#by material (plastic comparisons only)
ggplot(BC_ByMat_tibble[7:10,], aes(x = mat_comp, y = bc_score, fill = mat_comp)) +geom_col() +
  theme_rafa() + scale_fill_viridis_d()+
  coord_cartesian(ylim = c(0, 1)) +
  labs(title = "Community Dissimilarity between \n plastic and biodegradable bags", x = "Material Comparison",
       y = "Bray Curtis Dissimilarity", fill = "Comparisons")

#NMDS plots----
#extract scores from NMDS object
TC_NMDS_Scores <- as_tibble(scores(taxaCount_NMDS, display = "sites"))
TC_NMDS_Scores$site <- LPP_FullData$site
TC_NMDS_Scores$date <- LPP_FullData$date
TC_NMDS_Scores$material <- LPP_FullData$material

#plotting
ggplot(TC_NMDS_Scores, aes(x = NMDS1, y = NMDS2, fill = material)) +
  geom_hline(yintercept = 0, linetype = 5) + geom_vline(xintercept = 0, linetype = 5) +
  geom_point(shape = 21, size = 2.5) +
  facet_wrap(~date)+
  theme_rafa() +
  labs(title = "Bray Curtis NMDS of Community Matrices")+
  scale_fill_viridis_d()
```

```{r}
ggplot(water_quality, aes( x = sqrt(P_mean))) +geom_histogram(bins = 10)
ggplot(water_quality, aes( x = Nitrogen_mean)) +geom_histogram(bins = 15)
```

```{r}
#plot of first two NMDS dimensions, with ellipses for bag types
ordiplot(taxaCount_NMDS)
ordiellipse(taxaCount_NMDS,LPP_FullData$material, label = FALSE,
            col=viridis(n = 5),
            draw = "polygon", alpha=120) # adding ellipses to the plot, grouping by bag type
title("Similaity of Communities Across Bag Types")
legend("topright", title="Bag type",
       c("Biopolymer", "Cellulose" , "Cotton",  "Jute", "Plastic" ), fill=viridis(n = 5), horiz=FALSE, cex=.9) # adding a legend
```

```{r}
#prop plots of functional groups
FunGroups_props <- FunGroups %>% 
  select(!FnG_Richness & !fnG_Shannon) %>% 
  pivot_longer(cols = c("collector", "shredder", "scraper", "predator"), names_to = "fn_group", values_to = "counts") %>% 
  group_by(material, date, site) %>% 
  mutate(prop_counts = counts/sum(counts)) %>% 
  group_by(material, fn_group) %>% 
  mutate(avg_prop = mean(prop_counts),
         se_prop = sd(prop_counts)/sqrt(n())) %>% 
  summarise(avg_prop = unique(avg_prop),
            se_prop = unique(se_prop))

ggplot(FunGroups_props, aes(x = material, y = avg_prop, fill = fn_group)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = avg_prop -se_prop, ymax = avg_prop +se_prop), position = "dodge") +
  scale_fill_viridis_d(option = "D", begin = 0.2) +
  labs(title = "Distribution of Functional Groups across Bag types",
       y = "Avg. Proportion", caption = "error bars show Standard Error margins")+
  theme_rafa(base_size = 12)
```